A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2019-02-01 paper/submission/ZW10157-response-1.tex
adding Linda's last edits

The following information was added on lines 479-496:

Overall this solution is satisfactory, and meets all requirements for the new TBA beamline at the AWA.
A Power Extractor and Transfer Structure (PETS) will be located after the septum and quadrupoles in Fig. 7.
This device is putting the tightest constraints on the beam parameters.  
The aperture diameter is 17.6 mm and bunch length should be below 2 mm
to facilitate large power extraction [new ref here].

The optimization results shown here were used in part to meet all TBA requirements. 
After adjustments to include 3D rf field maps and CSR in the dipoles[new ref here] 
the max beamsizes at the PETS location were below 10 mm with a bunch length of rmsz=1.6 mm. 
Although the simulation results at the PETS location cannot yet be verified, since the beamline is not yet completely installed, 
some checks have been possible for the partially completed beamline.  
Simulations of the partial beamline have been done.
For example, beam images were taken on the YAG screen located downstream of the kicker and 
the extracted beam sizes at different kicker angles compared to simulation [new ref here].  
When it becomes possible to test the complete simulation solution, it is expected that there will not be perfect agreement with simulation;
for example there may be errors in the fields of beamline elements.  However, the simulation can be a tool to track down the source of
discrepancies, such as understanding the effect of the gun field on the beam symmetry.  
More importantly, the optimization results should provide significant improvement in efficiency for achieving a good operational regime.

>>>>

The following information was added on lines 401-404:

These structures are called Power Extractor and Transfer Structure (PETS), 
and will be located after the septum and quadrupoles in Fig.~7.
This device is putting the tightest constraints on the beam parameters.  
The aperture diameter is 17.6 mm and the bunch length should be below 2 mm
to facilitate large power extraction~[new ref here]

The following information was added on lines 457-475:

Overall this solution is satisfactory, and meets all requirements for the new TBA beamline at the AWA.

The TBA line under construction includes a kicker and septum combination to route one bunch train to its own PETS, 
while the other bunch train is undeflected and goes to a different PETS.  
The optimization results shown here start at the gun and go through to the first quadrupole after the septum (s3=19.4~(m)).  
These results were used as the basis for the full solution up to the PETS [new ref here].
After adjustments to include 3D rf field maps and CSR in the dipoles~[new ref here], 
the maximum beam sizes at the PETS location were below 10mm with a bunch length of rmsz = 1.6mm.  
Although the simulation results at the PETS location cannot yet be verified, 
since the beamline is not yet completely installed, some checks have been possible for the partially completed beamline. 
For example, beam images were taken on the YAG screen located downstream of the kicker, 
and it was found that the extracted beam sizes at different kicker angles 
were in good agreement with simulation results~[new ref here]. 
When it becomes possible to test the complete simulation solution, 
it is expected that there will not be perfect agreement with simulation; 
for example there may be errors in the fields of beamline elements. 
However, the simulation can be a tool to track down the source of discrepancies, 
such as understanding the effect of the gun field on the beam symmetry. 
More importantly, the optimization results should provide significant improvement 
in efficiency for achieving a good operational regime.

### 2019-02-01 paper/submission/prab.tex
adding Linda's last edits

In contrast to approaches that are tightly coupled to the optimization
algorithm, the range of possible applications is much wider.
Even in cases where the mathematical model of the forward solver is not known
exactly, fixed or on the real time measurements can be used to guide the
search for the Pareto optimal solutions.
Finally, combining a multi-objective optimization framework such as 
the one presented with practical experience in the field should expedite 
the decision making process in the design and operation of particle 
accelerators.

>>>>

In contrast to approaches that are tightly coupled to the optimization
algorithm, the range of possible applications is much wider.
Even in cases where the mathematical model of the forward solver is not known
exactly, fixed or real time measurements can be used to guide the
search for the Pareto optimal solutions.
Finally, combining a multi-objective optimization framework, such as 
the one presented, with practical experience in the field should expedite 
the decision making process in the design and operation of particle accelerators.

### 2019-02-01 paper/submission/theor-experiment.tex
adding Linda's last edits

To ensure that the optimizer works correctly, the benchmark
  problem (\ref{eqn:bench}) was solved.
To that end, we use a metric for comparing the quality of a Pareto front.
Given a point in the Pareto set, we compute the $m$ dimensional volume (for
  $m$ objectives) of the dominated space, relative to a chosen origin.
This is visualized for $2$ objectives in Fig.~\ref{fig:hypervolume}.
For further information and details of the implementation see~\cite{whbb:12}.
Figure~\ref{fig:pisa_bench} and the corresponding hypervolume values in
  Table~\ref{tbl:bench_rms_error} show expected convergence.
The reference Pareto front is clearly very well approximated.
It took a total of 1100 function evaluations to perform this computation.
The hypervolume of the reference solution ($0.6575$) for our benchmark was
  computed by sampling the solution provided in~\cite{hbwh:05}.

\begin{figure}
  \centering
    \includegraphics[width=0.6\linewidth]{figures/valid_front}
  \caption{Variator benchmark after $1100$ function evaluations using binary
           crossover and independent gene mutations (each gene mutates with
           probability $p=\frac{1}{2}$) on a population of $100$
           individuals.}
  \label{fig:pisa_bench}
\end{figure}

\begin{table}%[h!]
\begin{center}
  \caption{Convergence of benchmark problem with errors relative to
    hypervolume of sampled reference solution.}
  \label{tbl:bench_rms_error}
  \begin{tabular}{c|c|c}
	\cline{1-3}
    tot.\ function  & hypervolume & relative error\\
    evaluations    & & \\
	\cline{1-3}
    100  &  0.859753 & $3.076 \times 10^{-1}$ \\
    200  &  0.784943 & $1.938 \times 10^{-1}$ \\
    500  &  0.685183 & $4.210 \times 10^{-2}$ \\
    900  &  0.661898 & $6.689 \times 10^{-3}$ \\
    1100 &  0.657615 & $1.749 \times 10^{-4}$ \\
	\cline{1-3}
  \end{tabular}
\end{center}
\end{table}

Table~\ref{tbl:bench_rms_error} shows satisfactory
  convergence to the sampled reference Pareto front after 1000 (plus the
  additional 100 evaluations for the initial population) function evaluations.

\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next the optimization framework is applied to a high charge beam line
 at the Argonne Wakefield Accelerator (AWA) facility. 
The goal of this optimization is to produce beams of electrons that meet 
design specifications; this includes number of particles (charge), energy, and particle distribution (characterized by beam sizes and energy spread).
As shown in Fig.~\ref{awa-linac}, the installed portion of the the 
beam line consists of an rf photocathode gun, 
two solenoids, and six linear accelerating cavities
followed by four quadrupoles and a stripline kicker. 
The charge of interest, 40 nC, is needed for Two Beam 
Acceleration (TBA) 
experiments performed at AWA \cite{gai_power_jing_2012,JING201872}, 
which motivates this work. 
Prior experimental results were limited by beam size when the beam passed through small aperture 
wakefield structures located downstream. 
In an attempt to maximize charge transmission in upcoming experiments,
magnet strengths of the solenoids and quadrupoles leading into the 
TBA section of the beam line were optimized, shown in Fig.~\ref{awa-tba}.
The simulation model includes from the gun to the septum.
The optimization location is chosen as the  
entrance to the first quadrupole on the dog leg ($s_3$), see Fig.~\ref{awa-tba}.
Minimizing beam sizes here will enable capture and further focusing before space charge effects dominate the beam. 
This will also enable cleaner transport through downstream elements.
%\vspace{-1em}

@@

In the drifts and linac tanks, $dT=\num{1E-11}$ was sufficient. 
However, it was not acceptable near the quadrupoles. 
For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but discrepancies are seen in the transverse 
(rms$_x$ and $\epsilon_x$) for low fidelity results. This discrepancy is 
what led to the decision to adjust the time steps w.r.t. beam line elements. 
In the linac and drift sections $dT=\num{1E-11}$ was used. 
Near sensitive elements such as the quadrupoles, kicker, and septum, 
a time step of $dT=\num{1E-12}$ was used.
The resulting simulations are low fidelity in most places, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig.~\ref{tstep}.  Mid fidelity simulations used steps of  $dT=\num{1E-12}$ everywhere.
The average run time of each simulation with the adjusted time steps was 1.6~minutes.
In comparison, the mid fidelity simulation ran for 18~minutes.
Note, a smaller time step, \num{1E-13}, is always used in the gun where the 
beam has low energy and changing rapidly.
%A summary of the parameters used is listed in Table \ref{fidelity}.

\begin{figure}%[h]
	\centering
	\includegraphics[width=0.8\linewidth]{figures/timestep_comparison_paper}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

@@

\subsubsection{TBA Optimization Problem}
With computational and hyper parameters set, 
the full optimization problem of interest is explored.
The objectives (beam sizes and energy spread) are calculated at 
$s_3=19.4$~m, located downstream of the septum, see Fig.~\ref{awa-tba}. 
Given the longitudinal location of $s_3$ (unless otherwise noted), 
we define the objectives and input parameters as:
\begin{align}
\text{min}  \quad & \text{rms}_{x}, \quad \text{rms}_{y} \label{eq:awa:p1}\\
& \text{rms}_{px}, \quad \text{rms}_{py}, \label{eq:awa:p2}\\
& \text{rms}_{s}, \quad \text{dE} \label{eq:awa:p4} \\
\text{constraints} \quad & \text{rms}_x < 0.1 \, (\text{m}) |_{s=s_1}\label{eq:awa:c1}\\
\quad & \text{rms}_y < 0.1\, (m) |_{s=s_1}\, \label{eq:awa:c2}\\
\quad & |\text{rms}_y - \text{rms}_x | < 0.005 \, (\text{m}) |_{s=s_1}\label{eq:awa:c3}\\
\text{subject to} \quad & q = 40 \, (\text{nC}) \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64 \, \, (\text{MV/m}) \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24 \, \text{or} \, 25\,\, (\text{MV/m}) \\
\quad & R_x = R_y = 9 \, (\text{mm}) \label{eq:awa:firstdvar}\\
\quad & \phi_{\text{gun}} =-20^\circ \label{eq:awa:gphidvar}\\
\quad & \phi_{\text{linac}} =-20^\circ \label{eq:awa:lastdvar}
\end{align}

The first four objectives, parameters (\ref{eq:awa:p1}) to (\ref{eq:awa:p2}),
minimize the transverse ($rms_{x,y}$) beam size and transverse momentum ($rms_{px,py}$)
at the location of interest in the beam line ($s_3$, see Fig.~\ref{awa-tba}). 
Minimizing the beam size at this location is essential to 
to preventing loss of particles by scraping; 
which ensures better transmission through the wakefield structures downstream. 
Less divergence in the beam (lower transverse momentum spread) 
reduces growth of transverse beam size after the focal point (location of min beam size).
This reduces halo by ensuring the beam is not over focused through a hard waist.
The momentum spread is also critical to preventing large growth during transport. 
All of these factors help with transmission downstream.

The next two objectives in parameter (\ref{eq:awa:p4}) minimize the 
longitudinal beam size ($rms_s$), and energy spread (dE) at location $s_3$. 
This helps reduce the transverse beam size growth in bending elements.
A small bunch length ($rms_s$) is also critical to the goals of 
TBA experiments. The power generated in the wakefield structures 
designed for TBA is related to the bunch length \cite{JING201872,PETSeq}.

Eqs.~\ref{eq:awa:c1} to \ref{eq:awa:c3} 
define three constraints used to guide the algorithm.
However, it is important to not over-constrain the problem, which would prevent
the algorithm from converging.
The difference constraint, Eq.~\ref{eq:awa:c3}, is used to favor nearly round beams.
This prevents one dimension from becoming disproportionately large compared to the other.
At the AWA, there is some room in the beam pipe to allow the y dimension to grow, but round beams are preferred.

Equations (\ref{eq:awa:firstconstr}) to
(\ref{eq:awa:lastdvar}) define the charge, gun voltage, linac voltages, 
laser radius, gun phase, and linac cavity phases (in that order). 
These are parameters in the simulation 
that must be defined, but do not vary during the optimization.
The AWA design variables, objectives, and constraints in the OPAL input file as shown in
the repository.
%\vspace{0.2cm}

Design variables include the currents in two gun solenoids (IBF and IM), 
and four quadrupole strengths (KQ1-KQ4). The objectives include
beam size (transverse and longitudinal), transverse momentum, and energy spread as
defined in Eqs. (\ref{eq:awa:p1}) to (\ref{eq:awa:p4}). 
The location at the entrance of the kicker is $s_1=16.45$~meters, 
and the objectives are optimized at location $s_3=19.4$~meters. 
This is the entrance to the fifth quad in the beam line. 
This location is where the beam should be captured and focused through subsequent elements.

@@

With the time steps and hyper parameters set by the work in Section \ref{awa:subsection:test}, 
the full optimization problem described in \ref{awaproblem} was run for 200 generations.
The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers were in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
a large population size that would fit evenly on these resources was chosen. 
Again, the location of optimization is $s_3=19.4$ [m]. 
\begin{figure}
	\begin{center}		
		\includegraphics[width=0.65\textwidth]{figures/paperedit_pareto_front_Q5_xy_vs_pxpy}
	\end{center} 
		\caption{Pareto front comparing transverse beam sizes (rms$_{x,y}$) and transverse momentum (rms$_{px,py}$). The yellow star indicates the 
		point plotted in Fig.~\ref{fig:stat}}
	\label{fig:pareto1}
\end{figure}

As expected, the x dimension is impacted by the bending elements, and unable to reach 
the small beam sizes seen in the y dimension. This suggests objectives in the x 
dimension will drive design variable choices used during operations. 
However, it is still necessary to 
include the y dimension in the optimization. Early optimization tests showed the y dimension 
can easily grow out of control if it is not included in the objectives.
Those results are not shown here due to the unfeasible nature of the solutions 
(i.e. $rms_y$ larger than the beam pipe).
In the case of bunch length, there are not many options to choose from.

With these observations in mind, several beam parameters corresponding to
options on the Pareto Front in Fig.~\ref{fig:pareto1} were plotted and compared. 
A select result is shown in Fig.~\ref{fig:stat}.  
The maximum beam sizes are well below the beam pipe aperture limits as shown in Fig.~\ref{fig:stat}.
The solution is nearly round, which will increase changes of keeping the beam nearly round
as it travels to the last triplet in Fig.~\ref{awa-tba}.
Overall this solution is satisfactory, and meets all requirements for the new TBA beamline at the AWA.
A Power Extractor and Transfer Structure (PETS) will be located after the septum and quadrupoles in Fig.~\ref{awa-tba}.
This device is putting the tightest constraints on the beam parameters.  
The aperture diameter is \SI{17.6}{mm}, and bunch length should be below \SI{2}{mm}
to facilitate large power extraction~\cite{PETSeq}.

The optimization results shown here were used in part to meet all TBA requirements. 
After adjustments to include 3D rf field maps and CSR in the dipoles~\cite{neveuthesis}, 
the max beamsizes at the PETS location were below \SI{10}{mm} with a bunch length of $rms_z=\SI{1.6}{mm}$. 
Although the simulation results at the PETS location cannot yet be verified, since the beamline is not yet completely installed, 
some checks have been possible for the partially completed beamline.  
Simulations of the partial beamline have been done.
For example, beam images were taken on the YAG screen located downstream of the kicker and 
the extracted beam sizes at different kicker angles compared to simulation~\cite{neveuthesis}.  
When it becomes possible to test the complete simulation solution, it is expected that there will not be perfect agreement with simulation;
for example there may be errors in the fields of beamline elements.  However, the simulation can be a tool to track down the source of
discrepancies, such as understanding the effect of the gun field on the beam symmetry.  
More importantly, the optimization results should provide significant improvement in efficiency for achieving a good operational regime.

>>>>

To ensure that the optimizer works correctly, the benchmark
  problem (\ref{eqn:bench}) was solved.
To that end, we use a metric for comparing the quality of a Pareto front.
Given a point in the Pareto set, we compute the $m$ dimensional volume (for
  $m$ objectives) of the dominated space, relative to a chosen origin.
This is visualized for $2$ objectives in Fig.~\ref{fig:hypervolume}.
For further information and details of the implementation see~\cite{whbb:12}.
Figure~\ref{fig:pisa_bench} and the corresponding hypervolume values in
  Table~\ref{tbl:bench_rms_error} show expected convergence.
The reference Pareto front is clearly very well approximated.
It took a total of 1100 function evaluations to perform this computation.
The hypervolume of the reference solution ($0.6575$) for our benchmark was
 computed by sampling the solution provided in~\cite{hbwh:05}.
\begin{figure}
  \centering
    \includegraphics[width=0.6\linewidth]{figures/valid_front}
  \caption{Variator benchmark after $1100$ function evaluations using binary
           crossover and independent gene mutations (each gene mutates with
           probability $p=\frac{1}{2}$) on a population of $100$
           individuals.}
  \label{fig:pisa_bench}
\end{figure}
\begin{table}%[h!]
\begin{center}
  \caption{Convergence of benchmark problem with errors relative to
    hypervolume of sampled reference solution.}
  \label{tbl:bench_rms_error}
  \begin{tabular}{c|c|c}
	\cline{1-3}
    tot.\ function  & hypervolume & relative error\\
    evaluations    & & \\
	\cline{1-3}
    100  &  0.859753 & $3.076 \times 10^{-1}$ \\
    200  &  0.784943 & $1.938 \times 10^{-1}$ \\
    500  &  0.685183 & $4.210 \times 10^{-2}$ \\
    900  &  0.661898 & $6.689 \times 10^{-3}$ \\
    1100 &  0.657615 & $1.749 \times 10^{-4}$ \\
	\cline{1-3}
  \end{tabular}
\end{center}
\end{table}
Table~\ref{tbl:bench_rms_error} shows satisfactory
  convergence to the sampled reference Pareto front after 1000 (plus the
  additional 100 evaluations for the initial population) function evaluations.

\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next the optimization framework is applied to the high charge beam line
 at the Argonne Wakefield Accelerator (AWA) facility. 
The goal of this optimization is to produce beams of electrons that meet 
design specifications; this includes number of particles (charge), energy, 
and particle distribution (characterized by beam sizes and energy spread).
As shown in Fig.~\ref{awa-linac}, the installed portion of the the 
beam line consists of an rf photocathode gun, 
two solenoids, and six linear accelerating cavities
followed by four quadrupoles and a stripline kicker. 
The charge of interest, 40 nC, is needed for two beam acceleration (TBA) 
experiments performed at the AWA~\cite{gai_power_jing_2012,JING201872}, 
which motivates this work. 
Prior experimental results were limited by beam size when the beam passed through small aperture 
wakefield structures located downstream. 
In an attempt to maximize charge transmission in upcoming experiments,
magnet strengths of the solenoids and quadrupoles leading into the 
TBA section of the beam line were optimized, shown in Fig.~\ref{awa-tba}.
The simulation model includes from the gun to the septum.
The optimization location is chosen as the  
entrance to the first quadrupole on the dog leg ($s_3$), see Fig.~\ref{awa-tba}.
Minimizing beam sizes here will enable capture and further focusing before space charge effects dominate the beam. 
This will also enable cleaner transport through downstream elements.
%\vspace{-1em}

@@

In the drifts and linac tanks, $dT=\num{1E-11}$ was sufficient. 
However, it was not acceptable near the quadrupoles. 
For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but discrepancies are seen in the transverse 
(rms$_x$ and $\epsilon_x$) for low fidelity results. This discrepancy is 
what led to the decision to adjust the time steps w.r.t. beam line elements. 
In the linac and drift sections $dT=\num{1E-11}$ was used. 
Near sensitive elements such as the quadrupoles, kicker, and septum, 
a time step of $dT=\num{1E-12}$ was used.
The resulting simulations are low fidelity in most places, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig.~\ref{tstep}.  Mid fidelity simulations used steps of  $dT=\num{1E-12}$ everywhere.
The average run time of each simulation with the adjusted time steps was 1.6~minutes.
In comparison, the mid fidelity simulation ran for 18~minutes.
Note, a smaller time step, \num{1E-13}, is always used in the gun where the 
beam has low energy and is changing rapidly.
\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{figures/timestep_comparison_paper}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

@@

\subsubsection{TBA Optimization Problem}
With computational and hyper parameters set, 
the optimization problem of interest is explored.
The objectives (beam sizes and energy spread) are calculated at 
$s_3=19.4$~m, located downstream of the septum, see Fig.~\ref{awa-tba}. 
Given the longitudinal location of $s_3$ (unless otherwise noted), 
we define the objectives and input parameters as:
\begin{align}
\text{min}  \quad & \text{rms}_{x}, \quad \text{rms}_{y} \label{eq:awa:p1}\\
& \text{rms}_{px}, \quad \text{rms}_{py}, \label{eq:awa:p2}\\
& \text{rms}_{s}, \quad \text{dE} \label{eq:awa:p4} \\
\text{constraints} \quad & \text{rms}_x < 0.1 \, (\text{m}) |_{s=s_1}\label{eq:awa:c1}\\
\quad & \text{rms}_y < 0.1\, (m) |_{s=s_1}\, \label{eq:awa:c2}\\
\quad & |\text{rms}_y - \text{rms}_x | < 0.005 \, (\text{m}) |_{s=s_1}\label{eq:awa:c3}\\
\text{subject to} \quad & q = 40 \, (\text{nC}) \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64 \, \, (\text{MV/m}) \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24 \, \text{or} \, 25\,\, (\text{MV/m}) \\
\quad & R_x = R_y = 9 \, (\text{mm}) \label{eq:awa:firstdvar}\\
\quad & \phi_{\text{gun}} =-20^\circ \label{eq:awa:gphidvar}\\
\quad & \phi_{\text{linac}} =-20^\circ \label{eq:awa:lastdvar}
\end{align}

The first four objectives, parameters (\ref{eq:awa:p1}) to (\ref{eq:awa:p2}),
minimize the transverse ($rms_{x,y}$) beam size and transverse momentum ($rms_{px,py}$)
at the location of interest in the beam line ($s_3$, see Fig.~\ref{awa-tba}). 
Minimizing the beam size at this location is essential to 
to preventing loss of particles by scraping; 
which ensures better transmission through the wakefield structures downstream.
These structures are called Power Extractor and Transfer Structure (PETS), 
and will be located after the septum and quadrupoles in Fig.~\ref{awa-tba}.
This device is putting the tightest constraints on the beam parameters.  
The aperture diameter is \SI{17.6}{mm} and the bunch length should be below \SI{2}{mm}
to facilitate large power extraction~\cite{PETSeq}.   
Less divergence in the beam (lower transverse momentum spread) 
reduces growth of transverse beam size after the focal point (location of min beam size).
This reduces halo by ensuring the beam is not over focused through a hard waist.
The momentum spread is also critical to preventing large growth during transport. 
All of these factors help with transmission downstream.

The next two objectives in parameter (\ref{eq:awa:p4}) minimize the 
longitudinal beam size ($rms_s$), and energy spread (dE) at location $s_3$. 
This helps reduce the transverse beam size growth in bending elements.
A small bunch length ($rms_s$) is also critical to the goals of 
TBA experiments. The power generated in the wakefield structures 
designed for TBA is related to the bunch length \cite{JING201872,PETSeq}.
Eqs.~\ref{eq:awa:c1} to \ref{eq:awa:c3} 
define three constraints used to guide the algorithm.
However, it is important to not over-constrain the problem, which would prevent
the algorithm from converging.
The difference constraint, Eq.~\ref{eq:awa:c3}, is used to favor nearly round beams.
This prevents one dimension from becoming disproportionately large compared to the other.
At the AWA, there is some room in the beam pipe to allow the y dimension to grow, but round beams are preferred.
Equations (\ref{eq:awa:firstconstr}) to
(\ref{eq:awa:lastdvar}) define the charge, gun voltage, linac voltages, 
laser radius, gun phase, and linac cavity phases (in that order). 
These are parameters in the simulation that must be defined, but do not vary during the optimization.
For setup of the AWA design variables, objectives, and constraints in the OPAL input file, 
refer to the the repository above.

Design variables include the currents in two gun solenoids (IBF and IM), 
and four quadrupole strengths (KQ1-KQ4). The objectives include
beam size (transverse and longitudinal), transverse momentum, and energy spread as
defined in Eqs. (\ref{eq:awa:p1}) to (\ref{eq:awa:p4}). 
The location at the entrance of the kicker is $s_1=16.45$~(m), 
and the objectives are optimized at location $s_3=19.4$~(m). 
This is the entrance to the fifth quad in the beam line. 
This location is where the beam should be captured and focused through subsequent elements.

@@

With the time steps and hyper parameters set by the work in Section \ref{awa:subsection:test}, 
the optimization problem described in \ref{awaproblem} was run for 200 generations.
The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers were in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
a large population size that would fit evenly on these resources was chosen. 
Again, the location of optimization is $s_3=19.4$~(m). 
\begin{figure}
	\begin{center}		
		\includegraphics[width=0.65\textwidth]{figures/paperedit_pareto_front_Q5_xy_vs_pxpy}
	\end{center} 
		\caption{Pareto front comparing transverse beam sizes (rms$_{x,y}$) and transverse momentum (rms$_{px,py}$). The yellow star indicates the 
		point plotted in Fig.~\ref{fig:stat}}
	\label{fig:pareto1}
\end{figure}

As expected, the x dimension is impacted by the bending elements, and unable to reach 
the small beam sizes seen in the y dimension. This suggests objectives in the x 
dimension will drive design variable choices used during operations. 
However, it is still necessary to 
include the y dimension in the optimization. Early optimization tests showed the y dimension 
can easily grow out of control if it is not included in the objectives.
Those results are not shown here due to the unfeasible nature of the solutions 
(i.e. $rms_y$ larger than the beam pipe).
In the case of bunch length, there are not many options to choose from, as the phase was not varied.
With these observations in mind, several beam parameters corresponding to
options on the Pareto Front in Fig.~\ref{fig:pareto1} were plotted and compared. 
A select result is shown in Fig.~\ref{fig:stat}.  
The maximum beam sizes are well below the beam pipe aperture limits, also shown in Fig.~\ref{fig:stat}.
The solution is nearly round, which will increase chances of keeping the beam nearly round
as it travels to the last triplet in Fig.~\ref{awa-tba}.
Overall this solution is satisfactory, and meets all requirements for the new TBA beamline at the AWA.

The TBA line under construction includes a kicker and septum combination to route one bunch train to its own PETS, 
while the other bunch train is undeflected and goes to a different PETS.  
The optimization results shown here start at the gun and go through to the first quadrupole after the septum ($s_3=19.4$~(m)).  
These results were used as the basis for the full solution up to the PETS~\cite{neveuthesis}.
After adjustments to include 3D rf field maps and CSR in the dipoles~\cite{neveuthesis}, 
the maximum beam sizes at the PETS location were below 10mm with a bunch length of rmsz = 1.6mm.  
Although the simulation results at the PETS location cannot yet be verified, 
since the beamline is not yet completely installed, some checks have been possible for the partially completed beamline. 
For example, beam images were taken on the YAG screen located downstream of the kicker, 
and it was found that the extracted beam sizes at different kicker angles 
were in good agreement with simulation results~\cite{neveuthesis}. 
When it becomes possible to test the complete simulation solution, 
it is expected that there will not be perfect agreement with simulation; 
for example there may be errors in the fields of beamline elements. 
However, the simulation can be a tool to track down the source of discrepancies, 
such as understanding the effect of the gun field on the beam symmetry. 
More importantly, the optimization results should provide significant improvement 
in efficiency for achieving a good operational regime.
